Evolution of pre-copulatory and post-copulatory inbreeding behaviour and associated polyandry

### 2016-10-31 ms_pre_post.Rmd
Remove unnecessary comment
Last commit before getting to coauthor comments
Edit first paragraph of Introduction

---
title: "Evolution of pre-copulatory and post-copulatory inbreeding behaviour"
author: "A. Bradley Duthie et al."
date: "August 24, 2016"
output:
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
    reference_docx: docx_template.docx
  pdf_document:
    fig_caption: yes
  html_document: default
header-includes:
- \usepackage{natbib}
- \usepackage{lineno}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 2
csl: evolution.csl
bibliography: ms_pre_post.bib
---

@@

Inbreeding typically reduces the fitness of inbred offspring (termed 'inbreeding depression'), potentially driving evolution of inbreeding avoidance when inbreeding depression is strong [@Parker1979; @Parker2006; @Pusey1996; @Charlesworth1999; @Keller2002; @Charlesworth2009; @Szulkin2012]. For focal females, inbreeding avoidance might be acheived by avoiding related potential mates (pre-copulatory inbreeding avoidance), or by biasing fertilisation after mating has occurred so that related mates are less likely to sire offspring (post-copulatory inbreeding avoidance). Importantly, evolution of post-copulatory inbreeding avoidance requires that females be polyandrous, mating with multiple males during a single reproductive bout to ensure that post-copulatory choice can be expressed. As a reproductive strategy, polyandry thereby simultaneously allows females to choose multiple mates and creates the opportunity for further cryptic choice enacted through post-copulatory mechanisms. Polyandry has consequently been widely hypothesised to evolve as an adaptation to allow females to avoid inbreeding and therefore produce higher fitness offspring [@Zeh1997; @Jennions2000; @Tregenza2002; @Akcay2007]. However, how pre- and post-copulatory strategies of inbreeding avoidance evolve jointly, and how such evolution is predicted to feed back to affect evolution of polyandry, still remains largely unknown. While pre-copulatory inbreeding avoidance can facilitate evolution of polyandry under some restrictive conditions [@Duthie], the opportunity for post-copulatory choice might facilitate evolution of polyandry under a much wider range of conditions by allowing females to bias fertilisation toward non-kin.

@@

In our model, a cost free inbreeding strategy precluded a more costly complementary inbreeding strategy from persisting in a focal population (Figure 2), meaning that long-term persistence of both pre- and post-copulatory inbreeding avoidance might not be expected in populations given cost asymmetry. Direct costs therefore strongly affected the evolution of the costly phenotype(s), meaning that quantifying costs is necessary for predicting evolution of phenotypes associated with inbreeding strategies. Quantifying direct costs of phenotypes associated with polyandry and mate choice is empirically challenging [@Pomiankowski1987; @Kokko2003; @Reid2015]. @Pomiankowski1987 categorised four types of costs that are relevant to mate choice, which we suggest are also relevant costs for polyandry, and include costs of elevated risks associated with (1) predation or (2) disease transmission, and costs incurred through (3) loss of time or (4) depletion of energy. In our model, we interpreted a cost of polyandy as an elevated risk of predation as might be incurred while searching for or courting mates [e.g., @Rowe1988; @Rowe1994], a cost of pre-copulatory inbreeding strategy as a lost of time [i.e., risk of not finding a mate in time due to choosiness; e.g., @Kokko2013], and a cost of post-copulatory inbreeding strategy as depletion of energy. While these costs are biologically realistic, different costs will be relevant for different populations and be realised in different ways. For example, increased risk of disease transmission has been observed for highly polyandrous females [@Roberts2015a], but such a cost would more realistically be applied to the number of realised mates a female has rather than her tendency for engaging in polyandry. Future models will therefore need to carefully consider the magnitudes and realisation of direct costs to accurately predict evolution of inbreeding strategies in populations. <!--- add Tuni citation for depletion of energy? --->

>>>>

---
title: "Evolution of pre-copulatory and post-copulatory inbreeding behaviour and associated polyandry"
author: "A. Bradley Duthie et al."
date: "August 24, 2016"
output:
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
    reference_docx: docx_template.docx
  pdf_document:
    fig_caption: yes
  html_document: default
header-includes:
- \usepackage{natbib}
- \usepackage{lineno}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 2
csl: evolution.csl
bibliography: ms_pre_post.bib
---

@@

Inbreeding often greatly reduces the fitness of inbred offspring [termed 'inbreeding depression'; @Charlesworth1999; @Keller2002; @Charlesworth2009]. Such strong inbreeding depression is widely hypothesised to drive evolution of inbreeding avoidance, which can be enacted through multiple forms of mate choice in species with obligate biparental reproduction [@Parker1979; @Parker2006; @Pusey1996; @Szulkin2012]. From the perspective of a focal breeding female, inbreeding avoidance might be achieved by avoiding mating with related males (i.e., pre-copulatory inbreeding avoidance), or by biasing fertilisation towards unrelated males after mating has occurred, meaning that related males are less likely to sire offspring (i.e., post-copulatory inbreeding avoidance). Importantly, evolution of post-copulatory inbreeding avoidance requires that females are polyandrous, defined as mating with multiple males during a single reproductive bout. Further, such polyandry could itself evolve because it allows females to mate with additional unrelated males following initial mating with a relative, thereby facilitating pre-copulatory inbreeding avoidance even in the absence of any post-copulatory female choice or otherwise biased fertilisation (e.g., given a 'fair raffle' among sperm provided by feamles' mates). As a reproductive strategy, polyandry can thereby simultaneously allow females to mate with less closely related males and creates the opportunity for further inbreeding avoidance enacted through active post-copulatory choice. Polyandry has consequently been widely hypothesised to evolve as an adaptation to allow females to avoid inbreeding and thereby increase offspring fitness [@Zeh1997; @Jennions2000; @Tregenza2002; @Akcay2007]. However, despite such well-established hypotheses, no modelling investigates conditions under which pre-copulatory or post-copulatory inbreeding avoidance, or both strategies, are likely to evolve, or predicts how such evolution will feed back to affect underlying evolution of polyandry.

<!--- ------------------------------------------------------------------------- --->
<!--- ------------- LEFT OFF HERE --------------------------------------------- --->
<!--- ------------------------------------------------------------------------- --->

@@

In our model, a cost free inbreeding strategy precluded a more costly complementary inbreeding strategy from persisting in a focal population (Figure 2), meaning that long-term persistence of both pre- and post-copulatory inbreeding avoidance might not be expected in populations given cost asymmetry. Direct costs therefore strongly affected the evolution of the costly phenotype(s), meaning that quantifying costs is necessary for predicting evolution of phenotypes associated with inbreeding strategies. Quantifying direct costs of phenotypes associated with polyandry and mate choice is empirically challenging [@Pomiankowski1987; @Kokko2003; @Reid2015]. @Pomiankowski1987 categorised four types of costs that are relevant to mate choice, which we suggest are also relevant costs for polyandry, and include costs of elevated risks associated with (1) predation or (2) disease transmission, and costs incurred through (3) loss of time or (4) depletion of energy. In our model, we interpreted a cost of polyandy as an elevated risk of predation as might be incurred while searching for or courting mates [e.g., @Rowe1988; @Rowe1994], a cost of pre-copulatory inbreeding strategy as a lost of time [i.e., risk of not finding a mate in time due to choosiness; e.g., @Kokko2013], and a cost of post-copulatory inbreeding strategy as depletion of energy. While these costs are biologically realistic, different costs will be relevant for different populations and be realised in different ways. For example, increased risk of disease transmission has been observed for highly polyandrous females [@Roberts2015a], but such a cost would more realistically be applied to the number of realised mates a female has rather than her tendency for engaging in polyandry. Future models will therefore need to carefully consider the magnitudes and realisation of direct costs to accurately predict evolution of inbreeding strategies in populations.
